### 2022-12-20 README.Rmd
added a few details to main files

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
# Please put your title here to include it in the file below.
Title <- "Deconstructing Simple and Choice Decision and Movement Time Across the Spectrum of Schizophrenia Genetic Risk"
```

# srt

[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/brainworkup/srt/master?urlpath=rstudio)

This repository contains the data and code for our paper:

> Authors, (YYYY).
> *`r Title`*.
> Name of journal/book <https://doi.org/xxx/xxx>

Our pre-print is online here:

> Authors, (YYYY).
> *`r Title`*.
> Name of journal/book, Accessed `r format(Sys.Date(), "%d %b %Y")`.
> Online at <https://doi.org/xxx/xxx>

### How to cite

Please cite this compendium as:

> Authors, (`r format(Sys.Date(), "%Y")`).
> *Compendium of R code and data for `r Title`*.
> Accessed `r format(Sys.Date(), "%d %b %Y")`.
> Online at <https://doi.org/xxx/xxx>

## Contents

The **analysis** directory contains:

-   [:file_folder: paper](/analysis/paper): R Markdown source document for manuscript. Includes code to reproduce the figures and tables generated by the analysis. It also has a rendered version, `paper.docx`, suitable for reading (the code is replaced by figures and tables in this file)
-   [:file_folder: data](/analysis/data): Data used in the analysis.
-   [:file_folder: figures](/analysis/figures): Plots and other illustrations
-   [:file_folder: supplementary-materials](/analysis/supplementary-materials): Supplementary materials including notes and other documents prepared and collected during the analysis.

## How to run in your browser or download and run locally

This research compendium has been developed using the statistical programming language R.
To work with the compendium, you will need installed on your computer the [R software](https://cloud.r-project.org/) itself and optionally [RStudio Desktop](https://rstudio.com/products/rstudio/download/).

You can download the compendium as a zip from from this URL: [master.zip](/archive/master.zip).
After unzipping: - open the `.Rproj` file in RStudio - run `devtools::install()` to ensure you have the packages this analysis depends on (also listed in the [DESCRIPTION](/DESCRIPTION) file).
- finally, open `analysis/paper/paper.Rmd` and knit to produce the `paper.docx`, or run `rmarkdown::render("analysis/paper/paper.Rmd")` in the R console

### Licenses

**Text and figures :** [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** See the [DESCRIPTION](DESCRIPTION) file

**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse

### Contributions

We welcome contributions from everyone.
Before you get started, please see our [contributor guidelines](CONTRIBUTING.md).
Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md).
By participating in this project you agree to abide by its terms.