### 2022-12-20 analysis/paper/paper.qmd
added arxiv and elsevier quarto extensions

---
title: "Deconstructing Simple and Choice Decision and Movement Time Across the Spectrum of Schizophrenia Genetic Risk"
author:
  - Joey Trampush:
      correspondence: "yes"
      email: joey.trampush@med.usc.edu
      orcid: 0000-0003-0830-508X
      institute:
        - usc
        - nimh
  - Samantha Fleck:
      institute: usc
      orcid: 0000-0000-0000-0000
  - Dwight Dickinson:
      institute: nimh
      orcid: 0000-0003-1689-0559
  - Juan Pérez:
      orcid: 0000-0003-1689-0551
      institute:
        - name: Acme Corporation
  - Max Mustermann:
      orcid: 0000-0003-1689-0552
institute:
  - ksom:
      name: Keck School of Medicine of USC
      address: 23 Science Street, Eureka, Mississippi, USA
  - nimh: National Institute of Mental Health
title-block-published: "Last updated"  
date: now
date-format: long

format: 
  docx:
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
  pdf:
    cite-method: biblatex
  arxiv-pdf:
    keep-tex: true  
    linenumbers: false
    doublespacing: false
    runninghead: "Schizophrenia Reaction Time"
    cite-method: biblatex
  arxiv-html: default

execute:
  echo: true
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  schizophrenia; reaction time; keyword 3
highlights: |
  These are the highlights. 
---

```{r}
#| label: setup
#| include: false
library("papaja")
r_refs("references.bib")
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", echo = F, warning = F)
```

```{r}
#| label: analysis-preferences
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

<!-- With the following code you can access and display values from the yml header above. -->

Keywords: `r rmarkdown::metadata$keywords`

Highlights: `r rmarkdown::metadata$highlights`

<!-- The actual document text starts here: -->

# Introduction

Here is a citation [@Marwick2017]

# Background

# Methods

# Results

<!-- Here's some example analysis code: -->

```{r}
#| label: get-data
#| eval: false
# Note the path that we need to use to access our data files when rendering this document
my_data <- read.csv(here::here('analysis/data/raw_data/my_csv_file.csv'))
```

```{r}
#| label: fig-demo-plot
#| fig-cap: "A plot of random numbers"
plot(rnorm(10))
```

Figure @fig-demo-plot shows how we can have a caption and cross-reference for a plot. Note that figure label and cross-references must both be prefixed with `fig-`

```{r}
#| label: demo-inline-code
#| echo: false
x <- round(pi, 2)
```

Here is an example of inline code `r x` in the middle of a sentence.

# Discussion

# Conclusion

# Acknowledgements

<!-- The following line inserts a page break  -->

\newpage

# References

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

::: {#refs}
:::

\newpage

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r}
#| label: colophon
#| cache: false

# which R packages and versions?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```